---
title: "MRCKF Vingette"
author: "Arun Srinivasan (uus91@psu.edu)"
date: "8/16/2021"
output: html_document
---

### Introduction

We demonstrate the use of MRCKF through the following vignette. To illustrate this method, we conduct the following simulation of data under the linear model assumed in the MRCKF manuscript namely,

$$Y=XB + E$$
where $Y$ denotes the $n \times K$ response matrix, $X$ is the $n\times p$ covariate matrix, $B$ is the $p\times K$ coefficient and $E$ is the $n\times K$ response error matrix where $K$ denotes the number of responses of interest. 

Firstly, we must load the MRCKF function into R.

```{r, warning=F, message=F}
library(MRCKF)
```

### Controlled Variable Selection across Multiple Responses for Compositional Data


To ensure replicability, we fix a random seed for generating this simulated data.
```{r}
set.seed(123)
```

In this section, we specify the underlying parameter settings and generate data under linear model.

```{r, warning=F, message=F}
# Initialize Parameters
n = 100        # number of samples
p = 200        # number of features (without reference)
pstar = p + 1  #number of features (including reference)
K = 2          # number of responses
A = 3          # signal amplitude
sp = .05       # sparsity percentage
errorRho = .7  # error rho for E

# Generate covariance of W
sigma = matrix(0, nrow = pstar, ncol = pstar)
for(i in 1:pstar){
  for(j in 1:pstar){
    sigma[i,j] = .5^(abs(i-j))
  }
}

# generate W (latent log-normal data)
means = rep(1, pstar)
W = mvlognormal(n, means, Sigma = diag(sigma), R = sigma)
colnames(W) = paste("Taxa", 1:ncol(W))

# generate B (coefficient matrix)
Aset = seq(-A,A)
if(length(which(Aset==0))!=0){
  Aset = Aset[-which(Aset==0)]
}
B = matrix(0, nrow = p, ncol = K)
totalSig = p*K
nSig = floor(totalSig*sp)
signalEntries = sort(sample(1:totalSig, size = nSig, replace = F))
Bset = sample(Aset, nSig, replace = T)
B[signalEntries] = Bset


# generate error matrix E
errorCovariance = matrix(0,K, K)
for(i in 1:K){
  for(j in 1:K){
    errorCovariance[i,j] = errorRho^abs(i-j)
  }
}

E = matrix(0, nrow = n, ncol = K)
for(i in 1:n){
  ei = rnorm(K, mean = rep(0, K), errorCovariance)
  E[i,] = ei
}

# generate Z (compositional matrix)
Z = acomp(W)
Z = as.matrix(Z)

# generate X (alr-transformation matrix)
ref = Z[,pstar]
X = matrix(0, nrow = n, ncol = p)
for(t in 1:ncol(X)){
  X[,t] = log(Z[,t]/ref)
}

# generate Y (response matrix)
Y = X%*%B + E
colnames(Y) = paste("Response", 1:ncol(Y))
```


### Example Implementation

To inovke the MRCKF method, we call the MRCKF function with the default settings. We will target a $q=.2$ threshold.

```{r, warning=F, message=F, results = 'hide'}
q = .2         # targeted FDP threshold
result = MRCKF(W, Y, q)
```

As the coefficient matrix $B$ is large, we print out the first 10 rows of the selection matrices under both the MRCKF and MRCKF+ thresholds.

```{r}
# display the first 10 rows of the selection matrix under the MRCKF threshold
print(result$S[1:10,])

# display the first 10 rows of the selection matrix under the MRCKF+ threshold
print(result$Sp[1:10,])
```


Through this, we are able to compute the empirical FDP and Power by comparing the selected elements of the selection matrix to the indices of the non-zero entries of $B$.

```{r}
selectionSetS = which(result$S != 0)
selectionSetSp = which(result$Sp != 0)

truth = which(B != 0)

# compute selection statistics
fdrS = computeFDR(selectionSetS, truth)
powerS = computePower(selectionSetS, truth)
fdrSp = computeFDR(selectionSetSp, truth)
powerSp = computePower(selectionSetSp, truth)
```

```{r, echo=F}
paste("The estimated FDP under the MRCKF threshold is", fdrS)
paste("The estimated Power under the MRCKF threshold is", powerS)
paste("The estimated FDP under the MRCKF+ threshold is", fdrSp)
paste("The estimated Power under the MRCKF+ threshold is", powerSp)
```

### Single Response Setting

The MRCKF method can also be applied to setting with a single response. In this case, we assume the model

$$Y = X^Tb+\varepsilon$$

where $X$ denotes the alr-transformed data $(n \times p)$, $Y$ is the $n$-dimensional response, $b$ is the $p$-dimensional vector of coefficients, and $\varepsilon\sim N(0, \sigma^2I)$. This follows nearly identically to the previous example; however, the response $Y$ is a vector instead of a matrix.

```{r}
set.seed(123)
# Initialize Parameters
n = 100        # number of samples
p = 200        # number of features (without reference)
pstar = p + 1  #number of features (including reference)
K = 2          # number of responses
A = 3          # signal amplitude
sp = .05       # sparsity percentage

# Generate covariance of W
sigma = matrix(0, nrow = pstar, ncol = pstar)
for(i in 1:pstar){
  for(j in 1:pstar){
    sigma[i,j] = .5^(abs(i-j))
  }
}

# generate W (latent log-normal data)
means = rep(1, pstar)
W = mvlognormal(n, means, Sigma = diag(sigma), R = sigma)
colnames(W) = paste("Taxa", 1:ncol(W))

# generate B (coefficient matrix)
Aset = seq(-A,A)
if(length(which(Aset==0))!=0){
  Aset = Aset[-which(Aset==0)]
}
b = rep(0,p)
totalSig = p
nSig = floor(totalSig*sp)
signalEntries = sort(sample(1:totalSig, size = nSig, replace = F))
Bset = sample(Aset, nSig, replace = T)
b[signalEntries] = Bset

# generate Z (compositional matrix)
Z = acomp(W)
Z = as.matrix(Z)

# generate X (alr-transformation matrix)
ref = Z[,pstar]
X = matrix(0, nrow = n, ncol = p)
for(t in 1:ncol(X)){
  X[,t] = log(Z[,t]/ref)
}

# generate Y (single response) with sigma = 1
Y = X%*%b+rnorm(n)
```


The MRCKF function under the singular response is called in the exact same manner as the multiple response setting. When the response is a vector, the MRCKF function automatically defaults to the standard Model-X knockoff approach.
```{r}
q = .2         # targeted FDP threshold
resultSingle = MRCKF(W, Y, q)
```

The output of MRCKF is slightly different in the single response setting, as it will return the list of significant predictors by the corresponding column name of $W$. We are able to easily match these names and compute the empirical FDP and Power accordingly.

```{r}
selectionSetS = resultSingle$S
selectionSetSp = resultSingle$Sp

# match the selected taxa to column index
S = which(colnames(W) %in% selectionSetS)
Sp = which(colnames(W) %in% selectionSetSp)

# true signals
truth = which(b != 0)

# compute selection statistics
fdrS = computeFDR(S, truth)
powerS = computePower(S, truth)
fdrSp = computeFDR(Sp, truth)
powerSp = computePower(Sp, truth)
```

```{r, echo=F}
paste("The estimated FDP under the Single Response MRCKF threshold is", fdrS)
paste("The estimated Power under the Single Response MRCKF threshold is", powerS)
paste("The estimated FDP under the Single Response MRCKF+ threshold is", fdrSp)
paste("The estimated Power under the Single Response MRCKF+ threshold is", powerSp)
```
